### 2023-01-31 code/analysis_code/exploratory.qmd
Working with copies/mL dataset

---
title: "Exploratory"
format:
  html:
    theme: default
---
Library
```{r}
library(tidyverse)
library(here)
```

Load Data
```{r}
copies<- read_csv(here("data", "processed_data", "clean_data_date.csv")) #This is the clean_data.csv from copies_calculator but with date manually added in,
```

Make mini dataframes for each site
```{r}
copies2<- copies %>%
  select(date,copies_mL,site) %>%
drop_na()

bo<- copies2 %>% filter(site %in% "BO")
c2 <- copies2 %>% filter(site %in% "C2")
rd<- copies2 %>% filter(site %in% "RD")
summary(copies2)
```

```{r}
rd %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL)
  )
```

```{r}
c2 %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL)
  )
```

```{r}
bo %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL)
  )
```
```{r}
copies2 %>% ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site))+
  facet_wrap(~site)
```