### 2023-02-03 code/processing_code/dust_conc_processing.qmd
Transfer Part 1 from readme to Manuscript.qmd

---
title: "Dust Concentration"
format:
  html:
    theme: default
---
Library
```{r}
library(tidyverse)
library(here)
```

Load Data
```{r}
nrl<- read_csv(here("data","raw_data","nrl_conc.csv"))
```

Make new dataframes 
```{r}
dust_tot<- nrl %>%  #Dust data for all hours
  select(dust,date,hr)

dust_0<- dust_tot %>% #Dust data at t = 0
  filter(hr %in% "0")

dust_6<-dust_tot %>% #Dust data at t = 6
  filter(hr %in% "6")

dust_12<- dust_tot %>% #Dust data at t =12
  filter(hr %in% "12")

dust_18<- dust_tot %>% #Dust data at t = 18
  filter(hr %in% "18")

dust_sum<- aggregate(dust_tot$dust, by=list(date=dust_tot$date),FUN=sum) %>%
  rename("dust" = "x")
#dust_sum<- dust_sum %>%
 #left_join(dust_sum)
```

```{r}
dust_sum %>% ggplot() +geom_line(
  aes(x = date, 
      y = x)
)
```