### 2023-02-03 code/analysis_code/exploratory_dust.qmd
Added Exploratory dust qmd

---
title: "Exploratory Dust"
format:
  html:
    theme: default
---
Library
```{r}
library(here)
library(tidyverse)
```

Load Data
```{r}
dust_sum<-readRDS(file = here("data", "processed_data", "dust_sum.rds"))
```

Daily Dataframe
```{r}
daily_dust_sum<- dust_sum %>%
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

Plot Daily Dust Conc
```{r}
daily_dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = dust)
)
```

Plot Total Dust Conc
```{r}
dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = dust)
)
```