### 2023-02-09 code/analysis_code/exploratory_dust.qmd

---
title: "Exploratory Dust"
format:
  html:
    theme: default
---
Library
```{r}
library(here)
library(tidyverse)
```

Load Data that has just DUST CONC
```{r}
dust_sum<-readRDS(file = here("data", "processed_data", "dust_sum.rds")) #Dataframe with copies_mL and dust concentration for 2022
dust_0 <- readRDS(here("data", "processed_data", "dust_0.rds"))
dust_6 <- readRDS(here("data", "processed_data", "dust_6.rds"))
dust_12 <- readRDS(here("data", "processed_data", "dust_12.rds"))
dust_18 <- readRDS(here("data", "processed_data", "dust_18.rds"))

```

Daily Dataframe with Sum dust
```{r}
daily_dust_sum<- dust_sum %>%
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10
```

Plot Daily Dust Conc
```{r}
dust_sum_daily<- daily_dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = dust))

dust_sum_daily
```

Plot Total Dust Conc
```{r}
dust_sum_2022<- dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = dust))

dust_sum_2022
```

Save as PNG
```{r}
png(file = here("results","plots", "dust_sum_daily.png"))
dust_sum_daily
dev.off()

png(file = here("results","plots", "dust_sum_2022.png"))
dust_sum_2022
dev.off()
```