### 2023-02-09 code/analysis_code/exploratory_vp.qmd

---
title: "Vp Explore"
format:
  html:
    theme: default
---

Libraries
```{r}
library(tidyverse)
library(here)
```

Data
```{r}
vp<- read_csv(file = here("data", "processed_data", "vp_clean_copies_date.csv")) #This file had dates manually added
sq <-read_csv(here("data", "raw_data", "vp_qPCR_0.csv")) #Manually replaced NaN with 0 for no amplifcation 
```
Standard Curve
```{r}
std<- sq %>%
  filter(
    Fluor %in% "HEX",
    Content %in% "Std")

std %>%
  ggplot() + geom_point(
    aes(x = log10(SQ),
        y = Cq))

```

Make Site df
```{r}
rd6<- vp %>%
  filter(site %in% "RD")

c2<- vp %>%
  filter(site %in% "C2")
```

Plot Vp at C2
```{r}
c2_vp<- c2 %>% ggplot() + geom_line(
  aes(x = date,
      y = copies_mL),
  color = "purple") +
  geom_point(
    aes(x = date,
      y = copies_mL),
  color = "purple")+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vp at Canals") +
  theme(plot.title = element_text(hjust = 0.5))

c2_vp
```

Plot Vp at RD
```{r}
rd_vp<- rd6 %>% ggplot() + geom_line(
  aes(x = date,
      y = copies_mL),
  color = "red")+
  geom_point(
    aes(x = date,
      y = copies_mL),
  color = "red")+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vp at Gulf") +
  theme(plot.title = element_text(hjust = 0.5))

rd_vp
```

Save Vp C2/RD as PNG
```{r}
png(file = here("results", "plots", "vp_c2.png")) #Saving c2
c2_vp
dev.off()

png(file = here("results", "plots","vp_rd6.png")) #Saving RD
rd_vp
dev.off()
```