### 2023-02-09 code/processing_code/vp_copies_calculator.qmd

---
title: "Vp qPCR"
format:
  html:
    theme: default
---
*This dataset is Daily Vp qPCR*

Libraries
```{r}
library(tidyverse)
library(here)
```

Load data
```{r}
sq <-read_csv(here("data", "raw_data", "vp_qPCR_0.csv")) #Manually replaced NaN with 0 for no amplifcation 
```
Filter for HEX dye (samples). Cy5 is for IAC
```{r}
sq2<- sq %>% 
  select(filter_amt,SQ, Sample, Fluor) %>%
  filter(Fluor %in% "HEX")

```

SQ --> Copies/mL
```{r}
copies<- sq2 %>%
    mutate(
    #log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ)
    #sq = 10^log_sq, #Unlog 
    copies_rxn = SQ*15, #SQ*Reaction Volume
    copies_uL_1_10 = copies_rxn/1.2, #Copies/RXN / DNA Template Volume
    copies_uL_extract = copies_uL_1_10 *1, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
    copies_mL = copies_uL_extract*concentration_factor)%>%

  select(Sample,copies_mL)
```

Create Site Column from Sample Name 
```{r}
copies2<- copies %>%
  mutate(site = substr(copies$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
```

Drop NAs
```{r}
copies3<- copies2 %>% filter(!is.na(copies_mL)) %>% #Dropping NAs from STD Curve
  select(copies_mL,site, Sample)
```
*Dropping NAs got rid of BO completely*

Take the average of qPCR triplicates 
```{r}
copies4<- copies3 %>% group_by(site, Sample) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
  summarize_if(is.numeric, mean) %>%
  ungroup() #Ungroup to spit everything back out

```

Save as RDS and manually add date in excel
```{r}
write.csv(copies4, file= here("data","processed_data","vp_clean_copies.csv"))
```