### 2023-02-16 code/analysis_code/dust.qmd
Plots for symposium

---
title: "Exploratory Dust"
format:
  html:
    theme: default
---
Library
```{r}
library(here)
library(tidyverse)
library(lubridate)
```

Load Data that has just DUST CONC
```{r}
dust_sum<-readRDS(file = here("data", "processed_data", "dust_sum.rds")) #Dataframe with copies_mL and dust concentration for 2022
dust_0 <- readRDS(here("data", "processed_data", "dust_0.rds"))
dust_6 <- readRDS(here("data", "processed_data", "dust_6.rds"))
dust_12 <- readRDS(here("data", "processed_data", "dust_12.rds"))
dust_18 <- readRDS(here("data", "processed_data", "dust_18.rds"))

```

New Dataframes
```{r}
#Daily Dataframe with Sum dust
daily_dust_sum<- dust_sum %>%
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10

#Dataframe with just month dates
dust_month <- dust_sum %>%
  filter(date %in% as.Date(c("2022-02-16", "2022-03-09", "2022-04-20", "2022-05-11", "2022-06-14", "2022-07-07", "2022-08-09", "2022-09-13", "2022-10-18", "2022-11-15", "2022-12-13")))
```

Plot Daily Dust Conc
```{r}
dust_sum_daily<- daily_dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = dust))

dust_sum_daily
```

Plot Total Dust Conc
```{r}
dust_sum_2022<- dust_sum %>% ggplot() +geom_line(
 aes(x = date,
      y = dust),
  color = "#935d2e",
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Dust Aerosol Optical Density")+
  theme(
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1))+
  geom_vline(aes(xintercept =as.Date("2022-07-07")), linetype = "dotdash", color = "4c848c")+
  geom_vline(aes(xintercept =as.Date("2022-07-19")), linetype = "dotdash", color = "4c848c")+
  scale_x_date(date_breaks = "months" , date_labels = "%b") +

#Plot Month Sample Dates over
geom_point(
  data=dust_month,
  aes(x = date,
      y = dust),
  color = "4c848c",
  size = 3,
  shape = 4,
  stroke = 1)

dust_sum_2022
```

Save as PNG
```{r}
png(file = here("results","plots", "dust_sum_daily.png"))
dust_sum_daily
dev.off()

png(file = here("results","plots", "dust_sum_2022.png"))
dust_sum_2022
dev.off()
```