### 2023-02-16 code/analysis_code/vp.qmd

---
title: "Vp Explore"
format:
  html:
    theme: default
---

Libraries
```{r}
library(tidyverse)
library(here)
```

Data
```{r}
vp<- read_csv(file = here("data", "processed_data", "vp_clean_copies.csv")) #This file had dates manually added
#sq <-read_csv(here("data", "raw_data", "vp_qPCR.csv")) #Manually replaced NaN with 0 for no amplifcation 
```
Standard Curve
```{r}
#std<- sq %>%
  #filter(
    #Fluor %in% "HEX",
    #Content %in% "Std")
```

Make Site df
```{r}
rd6<- vp %>%
  filter(site %in% "RD")

c2<- vp %>%
  filter(site %in% "C2")
```

Plot Vp at C2
```{r}
c2_vp<- c2 %>% ggplot() + geom_line(
  aes(x = date,
      y = copies_mL),
  color = "purple") +
  geom_point(
    aes(x = date,
      y = copies_mL),
  color = "purple")+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vp at Canals") +
  theme(plot.title = element_text(hjust = 0.5))

c2_vp
```

Plot Vp at RD
```{r}
rd_vp<- rd6 %>% ggplot() + geom_line(
  aes(x = date,
      y = copies_mL),
  color = "red")+
  geom_point(
    aes(x = date,
      y = copies_mL),
  color = "red")+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vp at Gulf") +
  theme(plot.title = element_text(hjust = 0.5))

rd_vp
```
All Sites
```{r}
all<- vp %>% 
    mutate(site = recode(site, "BO" = "Blind Oso",
                              "C2" = "Canals",
                              "RD" = "Gulf")) %>%
  ggplot() + geom_line(
  aes(x = date,
      y = copies_mL,
      color = site)) +
  facet_wrap(~site)+
  geom_point(
    aes(x = date,
      y = copies_mL,
      color = site))+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vp at All Sites") +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)) +

 scale_color_manual( values = c("Blind Oso" = "darkgreen",
                                "Canals" = "darkred",
                                "Gulf" = "darkblue"))

all
```

Save Vp C2/RD as PNG
```{r}
png(file = here("results", "plots", "vp_all_sites.png")) #Saving c2
all
dev.off()

```