### 2023-02-16 code/wrangling_code/dust_conc_processing.qmd

---
title: "Dust Concentration"
format:
  html:
    theme: default
---
Library
```{r}
library(tidyverse)
library(here)
```

Load Data
```{r}
nrl<- read_csv(here("data","raw_data","nrl_conc.csv"))
copies<- read_csv(here("data", "processed_data", "clean_copies_date.csv"))
```

Make new dataframes  for Date and Dust for Exploratory 
```{r}
dust_tot<- nrl %>%  #Dust data for all hours
  select(dust,date,hr)

dust_0<- dust_tot %>% #Dust data at t = 0
  filter(hr %in% "0")

dust_6<-dust_tot %>% #Dust data at t = 6
  filter(hr %in% "6")

dust_12<- dust_tot %>% #Dust data at t =12
  filter(hr %in% "12")

dust_18<- dust_tot %>% #Dust data at t = 18
  filter(hr %in% "18")

dust_sum<- aggregate(dust_tot$dust, by=list(date=dust_tot$date),FUN=sum) %>% #could also use groupby
  rename("dust" = "x")
#dust_sum<- dust_sum %>%
 #left_join(dust_sum)
```

Make dataframe with Dates, Copies, and Concentration For Stats anaylsis
```{r}
copies2 <- copies %>% select(
  date, copies_mL, site)

dust_sum_copies<- dust_sum %>% left_join(copies2) #Takes df on right of pipe and joins to the df on the left of pipe

dust_t0 <- dust_0 %>% right_join(copies2)

dust_t6 <- dust_6 %>% right_join(copies2)

dust_t12<- dust_12 %>% right_join(copies2)

dust_t18<- dust_18 %>% right_join(copies2)

```

Save Dataframes as RDS for STATS (Copies and Dust)
```{r}
saveRDS(dust_sum_copies, file = here("data", "processed_data", "dust_sum_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_t0, file = here("data", "processed_data", "dust_0_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_t6, file = here("data", "processed_data", "dust_6_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_t12, file = here("data", "processed_data", "dust_12_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_t18, file = here("data", "processed_data", "dust_18_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

```

Save Dataframes as RDS for EXPLORE (Dust)
```{r}
saveRDS(dust_sum, file = here("data", "processed_data", "dust_sum.rds")) #Dataframe with dust concentration

saveRDS(dust_0, file = here("data", "processed_data", "dust_0.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_6, file = here("data", "processed_data", "dust_6.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_12, file = here("data", "processed_data", "dust_12.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_18, file = here("data", "processed_data", "dust_18.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
```