### 2023-02-24 code/analysis_code/dust.qmd

---
title: "Exploratory Dust"
format:
  html:
    theme: default
---
Library
```{r}
library(here)
library(tidyverse)
library(lubridate)
```

Load Data that has just dust concentrations
```{r}
dust_sum<-readRDS(file = here("data", "processed_data", "dust_conc","dust_sum.rds")) #Dataframe with copies_mL and dust concentration for 2022
#We will primary use the dust_sum data for exploratory analysis

dust_0 <- readRDS(here("data", "processed_data", "dust_conc", "dust_0.rds"))
dust_6 <- readRDS(here("data", "processed_data", "dust_conc", "dust_6.rds"))
dust_12 <- readRDS(here("data", "processed_data", "dust_conc", "dust_12.rds"))
dust_18 <- readRDS(here("data", "processed_data", "dust_conc", "dust_18.rds"))

```

New Data frame for Daily time series 
```{r}
#Daily Dataframe with Sum dust
daily_dust_sum<- dust_sum %>%
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10
```

Plot Daily Dust Concentration over Daily Time Series 
```{r}
dust_sum_daily<- daily_dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = dust))

dust_sum_daily
```

Save as PNG
```{r}
png(file = here("results","plots", "dust_sum_daily.png"))
dust_sum_daily
dev.off()
```