### 2023-02-24 code/analysis_code/dust_and_copies.qmd

---
title: "Dust_Copies_Plot"
format:
  html:
    theme: default
---

Library
```{r}
library(here)
library(tidyverse)
library(ggpubr)
```

Load Data that has just DUST CONC
```{r}
#Summed Dust Concentration per day
dust_sum<-readRDS(file = here("data", "processed_data", "dust_copies", "dust_copies_sum.rds")) 
copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) 
```

Daily Data frame with Summed dust by day
```{r}
daily_dust_sum<- dust_sum %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10. No sampling occured then.

copies_d <- copies %>%
  filter(Type %in% "Daily") %>%
  mutate(site = recode(site, "BO" = "Blind Oso",
                             "C2" = "Canals",
                            "RD" = "Gulf"))

```

Plot Daily Dust Concentration
```{r}
dust_fig<- daily_dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = dust),
  color = "#935d2e",
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Dust Aerosol Optical Density")+
  theme(
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 10))

dust_fig
```

Total Daily Vibrio
```{r}
tot_vib<- copies_d %>% ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site),
  size = 1) +
  theme_bw()+
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at all Sites",
    color = "  Site") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20),  #center title
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 10),
    legend.position = c(0.1, 0.8),
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
  scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

tot_vib
```

Combine Dust and Copies
```{r, fig.height = 4, fig.width= 4}
fig<- ggarrange(tot_vib, dust_fig + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")

fig
```

Save as PNG
```{r}
png(file = here("results", "plots", "vib_dust_daily.png"))
fig
dev.off()
```