### 2023-02-24 code/wrangling_code/dust_conc_processing.qmd

---
title: "Dust Concentration"
format:
  html:
    theme: default
---
This data set is sourced from the Naval Research Lab. Aerosol optical thickness (AOT) was measures at 4 time points per day (t = 0,6,12,18).

Library
```{r}
library(tidyverse)
library(here)
```

Load Data
```{r}
nrl<- read_csv(here("data","raw_data","nrl_conc.csv")) #Dust concentration data set
copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) #Vibrio counts data set from qPCR
```

Make new data frames for each time point as well as sum AOT per day
```{r}
dust_tot<- nrl %>%  #Dust data for all hours
  select(dust,date,hr)

dust_0<- dust_tot %>% #Dust data at t = 0
  filter(hr %in% "0")

dust_6<-dust_tot %>% #Dust data at t = 6
  filter(hr %in% "6")

dust_12<- dust_tot %>% #Dust data at t =12
  filter(hr %in% "12")

dust_18<- dust_tot %>% #Dust data at t = 18
  filter(hr %in% "18")

dust_sum<- aggregate(dust_tot$dust, by=list(date=dust_tot$date),FUN=sum) %>% #Sum of AOT by day
  rename("dust" = "x")

```

Make dataframe with Dates, Copies/mL, and Dust Concentration For Stats analysis. Here we are combining dust data with copies/mL data. 
```{r}
#Condense copies dateset
copies2 <- copies %>% select(
  date, copies_mL, site)

#Takes df on right of pipe and joins to the df on the left of pipe 
dust_copies_sum<- dust_sum %>% left_join(copies2) #Dust at all time points
dust_copies_0 <- dust_0 %>% right_join(copies2) #T0
dust_copies_6 <- dust_6 %>% right_join(copies2) #T6
dust_copies_12<- dust_12 %>% right_join(copies2) #T12
dust_copies_18<- dust_18 %>% right_join(copies2) #T18
```

Save **dust/copies** dataframes as RDS for Stats
```{r}
saveRDS(dust_copies_sum, file = here("data", "processed_data", "dust_copies", "dust_copies_sum.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_copies_0, file = here("data", "processed_data",  "dust_copies","dust_copies_0.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_copies_6, file = here("data", "processed_data", "dust_copies", "dust_copies_6.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_copies_12, file = here("data", "processed_data",  "dust_copies","dust_copies_12.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_copies_18, file = here("data", "processed_data",  "dust_copies","dust_copies_18.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

```

Save **Dust Concentration** data frames as RDS (good for Dust plots)
```{r}
saveRDS(dust_sum, file = here("data", "processed_data", "dust_conc", "dust_sum.rds")) #Dataframe with dust concentration

saveRDS(dust_0, file = here("data", "processed_data","dust_conc", "dust_0.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_6, file = here("data", "processed_data","dust_conc", "dust_6.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_12, file = here("data", "processed_data","dust_conc", "dust_12.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_18, file = here("data", "processed_data","dust_conc", "dust_18.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
```