### 2023-02-24 code/wrangling_code/total_copies_calculator.qmd

---
title: "Copies/mL"
format:
  html:
    theme: default
---
*This script is for cleaning qPCR results on Total Vibrio enumeration. It takes Cq value outputs from the CFX (qPCR machine) and computes it to copies of gene targets of Vibrio. The data set contains a daily and monthly time series. We will be examining daily.*

Libraries
```{r}
library(tidyverse)
library(here)
```

Load data
```{r}
data <-read_csv(here("data", "raw_data", "copies_master.csv"))
```

Select and Filter for Daily and Monthly Time Series 
```{r}
daily<- data %>%
select(Type,Sample, Cq) %>%
  filter(Type %in% "Daily")

monthly<- data %>%
select(Type,Sample, Cq) %>%
  filter(Type %in% "Monthly")
```

Create Site Column from Sample Name 
```{r}
site_daily<- daily %>%
  mutate(site = substr(daily$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)

site_monthly<- monthly %>%
  mutate(site = substr(monthly$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
```

Add Dilution Column and Filter column. The dilution column applies if samples were diluted prior to qPCR. For example, if I diluted a sample 1:10, my dilution column would denote "10".  The filter column pertains to the amount of water sample (mL) that was concentrated onto a water filter prior to qPCR. 
```{r}
dilution_daily<- site_daily %>%
  mutate(dilution = case_when(
                          Sample == "BO7"~"10", #BO site was diluted 1:10
                          Sample == "BO8"~"10",
                          Sample == "BO9"~"10",
                          Sample == "BO11"~"10",
                          Sample == "BO12"~"10",
                          Sample == "BO13"~"10",
                          Sample == "BO14"~"10",
                          Sample == "BO15"~"10",
                          Sample == "BO16REDO"~"1",
                          Sample == "BO17"~"10",
                          Sample == "BO18"~"10",
                          Sample == "BO19"~"10",
                              site  == "C2"~ "1", #Canals and Gulf were not diluted
                              site == "RD"~"1"), 
        dilution = as.numeric(dilution))

dilution_monthly<- site_monthly %>%
  mutate(dilution = case_when(
                              site == "BO"~ "1",
                              site  == "C2"~ "1",
                              site == "RD"~"1"), 
        dilution = as.numeric(dilution))

filter_daily<- dilution_daily %>%
   mutate(filter_amt = case_when(Sample == "BO_7"~"100", #100mL concnetrated onto filter
                          Sample == "BO7"~"100",
                          Sample == "BO8"~"100",
                          Sample == "BO9"~"100",
                          Sample == "BO11"~"100",
                          Sample == "BO12"~"50",
                          Sample == "BO13"~"50",
                          Sample == "BO14"~"50",
                          Sample == "BO15"~"50",
                          Sample == "BO16REDO"~"30", #30 mL concentrated onto filter 
                          Sample == "BO17"~"30",
                          Sample == "BO18"~"30",
                          Sample == "BO19"~"30",
                              site  == "C2"~ "100",
                              site == "RD"~"100"),
        filter_amt = as.numeric(filter_amt))

filter_monthly<- dilution_monthly %>%
   mutate(filter_amt = case_when(Sample == "BO_7"~"100",
                          Sample == "BO_FEB"~"100",
                          Sample == "BO_MAR"~"100",
                          Sample == "BO_APR"~"100",
                          Sample == "BO_MAY_11"~"100",
                          Sample == "BO_MAY_20"~"100",
                          Sample == "BO_MAY_24"~"100",
                          Sample == "BO_JUN"~"100",
                          Sample == "BO_AUG"~"30",
                          Sample == "BO_SEP"~"45",
                          Sample == "BO_OCT"~"60",
                          Sample == "BO_NOV"~"30",
                          Sample == "BO_DEC"~"60",
                              site  == "C2"~ "100",
                              site == "RD"~"100"),
        filter_amt = as.numeric(filter_amt)) 
```

Go from CQ --> Copies/mL
```{r}
copies_daily<- filter_daily %>%
    mutate(
    log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ):  This comes from running a standard curve during qPCR. 
    sq = 10^log_sq, #Unlog 
    copies_rxn = sq*10, #SQ*Reaction Volume: Each qPCR reaction was 10µL
    copies_uL_1_10 = copies_rxn/2.5, #Copies/RXN / DNA Template Volume: Each 10µL reaction had 2.5µL of DNA added. 
    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
    copies_mL = copies_uL_extract*concentration_factor) %>%

  select(Sample,copies_mL,site,Type)

copies_monthly<- filter_monthly %>%
    mutate(
    log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ)
    sq = 10^log_sq, #Unlog 
    copies_rxn = sq*10, #SQ*Reaction Volume
    copies_uL_1_10 = copies_rxn/2.5, #Copies/RXN / DNA Template Volume
    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
    copies_mL = copies_uL_extract*concentration_factor) %>%

 select(Sample,copies_mL,Type, site)

```

Some samples did not amplify: Replace NaN with 0 for copies/mL
```{r}
replace_daily<- copies_daily %>% 
 mutate(copies_mL = ifelse(is.na(copies_mL), 0, copies_mL)) %>%
filter(!is.na(Sample))   #Dropping NAs from STD Curve

replace_monthly<- copies_monthly %>% 
 mutate(copies_mL = ifelse(is.na(copies_mL), 0, copies_mL)) %>%
filter(!is.na(Sample))   #Dropping NAs from STD Curve 
```

Date Column: This adds a date correpsonding the a number in the sample name. 
```{r}
date_daily<- replace_daily %>%
  mutate(day = substr(replace_daily$Sample,3,4)) %>% #Creates new column `day` 
  mutate(date = case_when(day == "7"~"2022-07-07", #Makes date based off of day
                          day == "7R"~"2022-07-07",
                         day ==  "8"~"2022-07-08",
                         day ==  "9"~"2022-07-09",
                         day ==  "11"~"2022-07-11",
                         day ==  "12"~"2022-07-12",
                         day ==  "13"~"2022-07-13",
                         day ==  "14"~"2022-07-14",
                         day ==  "15"~"2022-07-15",
                         day ==  "16"~"2022-07-16",
                         day ==  "17"~"2022-07-17",
                         day ==  "18"~"2022-07-18",
                         day ==  "19"~"2022-07-19")) %>%
  select(copies_mL,site,date,Type)

date_monthly<- replace_monthly %>%
  mutate(month = substr(replace_monthly$Sample,4,9)) %>% #Creates new column `month` 
  mutate(date = case_when(month == "FEB"~"2022-02-16", #Makes date based off of day
                         month ==  "MAR"~"2022-03-09",
                         month ==  "APR"~"2022-04-20",
                         month ==  "MAY_11"~"2022-05-11",
                         month ==  "MAY_20"~"2022-05-20",
                         month ==  "MAY_24"~"2022-05-24",
                         month ==  "JUN"~"2022-06-14",
                         month ==  "AUG"~"2022-08-09",
                         month ==  "SEP"~"2022-09-13",
                         month ==  "OCT"~"2022-10-18",
                         month ==  "NOV"~"2022-11-15",
                         month ==  "DEC"~"2022-12-14")) %>%
  select(copies_mL,site,date,Type)

```

Each sample (Date) was done in triplicate for qPCR: We need to take the average of qPCR triplicates 
```{r}
final_daily<- date_daily %>% group_by(date,site,Type) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
  summarize_if(is.numeric, mean) %>%
  ungroup() #Ungroup to spit everything back out

final_monthly<- date_monthly %>% group_by(date,site,Type) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
 summarize_if(is.numeric, mean) %>%
  ungroup() #Ungroup to spit everything back out
```

Combine Daily and Monthly data sets back togthether. 
```{r}
copies_final<- full_join(final_monthly,final_daily)
summary(copies_final)
str(copies_final)
```

Save as RDS and CSV
```{r}
#saveRDS(copies_final, file= here("data","processed_data","total_clean_copies.rds"))
write.csv(copies_final,file= here("data","processed_data","total_clean_copies.csv"))

```