### 2023-02-24 data/raw_data/rawdata_README.Rmd

This folder contains dust concentrations in AOT from the Naval Reserach lab, environmental parameters (temperature and salinity) that were collected during the summer at the three sites, and  qPCR results for Total Vibrio.