### 2023-03-14 code/analysis_code/dust/daily_dusts.qmd
Cleaned up code

---
title: "New Dust and Copies"
format:
  html:
    theme: default
---
**This code is useful for looking at variations of daily dust trends. For example, when running statistical analysis, it may be best to look at dust concentration at a specific hour, or the sum per day, or the average. We will be working with a sum that is 24hr prior to sample collection. This utilizes a data set that has the variations of dust trends **
Libs
```{r}
library(here)
library(tidyverse)
```

Data
```{r}
dust<- read_csv(here("data", "processed_data", "dust_master.csv")) 
```
**There are a few columns that we do not need or that contain NaN. So let's remove those.**

Clean
```{r}
dust<- dust %>%
  select(date,t1,t7,t13,t19,tsum,t7sum) %>%
  filter(!is.na(dust$date))
```

Plot T7sum
```{r}
dust %>% ggplot()+ geom_line(
  aes(x = date,
      y = t7sum))+
  labs(title = "Dust Concentration over Daily Time Series",
       x = "Date",
       y = "Dust Concentration (AOD)")+
  theme_bw()
```