### 2023-03-14 data/processed_data/processed_data_README.Rmd

---

This folder contains processed data that has been cleaned from code in the `wrangling` folder.\
(1) `total_clean_copies`: Vibrio enumeration (as copies per mL) across time series.\
(2) `dust_clean.rds`: Cleaned dust data
(3) `ysi_2.csv`: Environmental parameter data
(4) `tx_master.csv` which is a compilation of all relevant data. 
(5)`dust_master.csv`: Manually created dust data examining different time points, sums, avereages (helpful for statistics)