### 2023-03-16 code/analysis_code/stats/dust_copies_stats.qmd
stats

---
title: "env_and_copies_stats"
format:
  html:
    theme: default
---
### Libraries
```{r}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
library(dotwhisker) #Fpr visualization
```

# LOADING AND CLEANING DATA

### Data
```{r}
dust0<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
data0<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data
```

### Clean and Trim
```{r}
dust<- dust0 %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))

data<- data0 %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

### Combine into one big data set
```{r}
master<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.

master<- master %>%
  select(copies_mL, t7sum, site, date) 
```

### Make Site DF
```{r}
bo<- master %>%
  filter(site %in% "Blind Oso"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06")))

c2<- master %>%
  filter(site %in% "Canals"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06")))

rd<- master %>%
  filter(site %in% "Gulf"| date %in% c(as.Date("2022-07-05"), as.Date("2022-07-06")))
```

#DUST

**From previous literature, we know that Vibrio have responded to dust 24-48 hours after input. This can make analysis tricky, as the dust concentration of a specific day does not necessarily correlate with the Vibrio copies of that same date. We can use cross-correlation analysis to assess a lag in a time series and see where this lag is most significant.**

### Cross-correlation:
```{r}
#Blind Oso
bo_ccf<- bo %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
ccf(bo_ccf$t7sum, bo_ccf$copies_mL) 
title("Blind Oso DustxCopies CCA", line =1)

#Canals
c2_ccf<- c2 %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
ccf(c2_ccf$t7sum, c2_ccf$copies_mL)
title("Canals DustxCopies CCA", line=1)

#Gulf
rd_ccf<- rd %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
ccf(rd_ccf$t7sum, rd_ccf$copies_mL) 
title("Gulf DustxCopies CCA", line =1)
```
**A negative Lag value indicates that x is a predictor of y, or in this case, dust is a predictor of copies_mL- so we will be looking at these negative values. A positive ACF indicates a positive relationship between the two variables, where a negative ACF indicates a negative relationship. According to the ccf plots, BO has a significant lag relationship a value of -1. In other words, dust occurs, and one day later, we see a response in copies_mL. C2 has a lag at -2, and RD at -1 **

## Gulf Site Dust

### Adding a 2 day lag in dust:
```{r}
rd2<- rd%>% mutate(
  lag =Lag(rd$t7sum, shift  = 2)) #Make lag from ccf
```

### Create a linear Model: 
```{r}
lm_spec<- linear_reg()
lm_fit<- fit(lm_spec, copies_mL~lag, data = rd2)
tidy(lm_fit)
```

### Plot RD T7 Sum:
```{r}
rd2 %>% ggplot() + geom_point(
  aes(x = lag, 
      y = copies_mL)) +
  geom_smooth(aes(x = lag,
                  y = copies_mL),
              method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (RD+2)")
```
## BO Site Dust

### Adding a 2 day lag in dust:
```{r}
bo2<- bo%>% mutate(
  lag =Lag(bo$t7sum, shift  = 1)) #Make lag from ccf
```

### Create a linear Model: 
```{r}
lm_spec<- linear_reg()
lm_fit<- fit(lm_spec, copies_mL~lag, data = bo2)
tidy(lm_fit)
```

### Plot BO T7 Sum:
```{r}
bo2 %>% ggplot() + geom_point(
  aes(x = lag, 
      y = copies_mL)) +
  geom_smooth(aes(x = lag,
                  y = copies_mL),
              method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (BO+1)")
```
## C2 Site Dust

### Adding a 2 day lag in dust:
```{r}
c22<- c2%>% mutate(
  lag =Lag(c2$t7sum, shift  = 2)) #Make lag from ccf
```

### Create a linear Model: 
```{r}
lm_spec<- linear_reg()
lm_fit<- fit(lm_spec, copies_mL~lag, data = c22)
tidy(lm_fit)
```

### Plot C2 T7 Sum:
```{r}
c22 %>% ggplot() + geom_point(
  aes(x = lag, 
      y = copies_mL)) +
  geom_smooth(aes(x = lag,
                  y = copies_mL),
              method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (C2+2)")
```

_______________________
All Variables

New data set
```{r}
master2<- dust0 %>% full_join(data) %>%
  filter(hr_cst %in% "01") %>%
  select(!c(t1, t7, t13, t19, hr_cst, tsum, t7sum, ...1, SUM, dust,  Type, date )) %>%
  na.omit()
```
```{r}
bo_a<- master2 %>%
  filter(site %in% "Blind Oso") %>%
  select(!site)

c2_a<- master2 %>%
  filter(site %in% "Canals")%>%
  select(!site)

rd_a<- master2 %>%
  filter(site %in% "Gulf")%>%
  select(!site)
```

## BO: Using all Variables vs. Growth
```{r}
#Create Recipe for Growth and all other parameters
growth_recipe_bo<- recipe(copies_mL ~ ., data = bo_a)

#Set up linear model
lm_mod_all<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod_all) %>%
  add_recipe(growth_recipe_bo)

#Use workflow to fit model to train data set
growth_fit_bo<- Growth_wflow %>%
  fit(data = bo_a)

#View as Tibble 
growth_fit_bo %>%
  extract_fit_parsnip() %>%
  tidy()
```
Plot
```{r}
tidy(growth_fit_bo) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```

## C2: Using all Variables vs. Growth
```{r}
#Create Recipe for Growth and all other parameters
growth_recipe_c2<- recipe(copies_mL ~ ., data = c2_a)

#Set up linear model
lm_mod_all<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod_all) %>%
  add_recipe(growth_recipe_c2)

#Use workflow to fit model to train data set
growth_fit_c2<- Growth_wflow %>%
  fit(data = c2_a)

#View as Tibble 
growth_fit_c2 %>%
  extract_fit_parsnip() %>%
  tidy()
```
Plot
```{r}
tidy(growth_fit_c2) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```
## RD: Using all Variables vs. Growth
```{r}
#Create Recipe for Growth and all other parameters
growth_recipe_rd<- recipe(copies_mL ~ ., data = rd_a)

#Set up linear model
lm_mod_all<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod_all) %>%
  add_recipe(growth_recipe_rd)

#Use workflow to fit model to train data set
growth_fit_rd<- Growth_wflow %>%
  fit(data = rd_a)

#View as Tibble 
growth_fit_rd %>%
  extract_fit_parsnip() %>%
  tidy()
```
Plot
```{r}
tidy(growth_fit_rd) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```