### 2023-03-16 data/raw_data/rawdata_README.Rmd

This folder contains two raw data sets:\
(1) `dust.csv`:dust concentrations in AOT from the Naval Research lab. Concentrations (AOT) were collected at 4 time points (UTC): T0,T6,T12, and T18.\
(2) `copies_master.csv`:  qPCR results for Total Vibrio enumeration.