### 2023-03-20 code/analysis_code/stats/dust_copies_stats.qmd
stats

---
title: "env_and_copies_stats"
format:
  html:
    theme: default
---

### Libraries

```{r}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
library(dotwhisker) #Fpr visualization
```

# LOADING AND CLEANING DATA

### Data

```{r}
dust0<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
data0<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data
```

### Clean and Trim

```{r}
dust<- dust0 %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))

data<- data0 %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

### Combine into one big data set

```{r}
master<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.

master<- master %>%
  select(copies_mL, t7sum, site, date) 
```

# Check Distributions

### Copies per mL

```{r}

#Normal
master %>% 
  ggplot() + geom_histogram(
    aes(
      x = copies_mL)) +
  labs(title = "Non-Log")+ 
  facet_wrap(~site)

#Log-Transformed
master %>% 
  ggplot() + geom_histogram(
    aes(
      x = log(copies_mL))) +
  labs(title = "LOG-TRANS")+ 
  facet_wrap(~site)
```

**We will use the log-transformed data as it has a \~ normal distribution.**

## Log-transform Copies

```{r}
#master <- 
 # master %>%
 # mutate(copies_mL = log(master$copies_mL))
```

### Make Site DF

```{r}
bo<- master %>%
  filter(site %in% "Blind Oso"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06")))

c2<- master %>%
  filter(site %in% "Canals"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06")))

rd<- master %>%
  filter(site %in% "Gulf"| date %in% c(as.Date("2022-07-05"), as.Date("2022-07-06")))
```

# Uni-variate Models

**From previous literature, we know that Vibrio have responded to dust 24-48 hours after input. This can make analysis tricky, as the dust concentration of a specific day does not necessarily correlate with the Vibrio copies of that same date. We can use cross-correlation analysis to assess a lag in a time series and see where this lag is most significant.**

## Cross-correlation:

```{r}
#Blind Oso
bo_ccf<- bo %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
ccf(bo_ccf$t7sum, bo_ccf$copies_mL) 
title("Blind Oso DustxCopies CCA", line =1)

#Canals
c2_ccf<- c2 %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
ccf(c2_ccf$t7sum, c2_ccf$copies_mL)
title("Canals DustxCopies CCA", line=1)

#Gulf
rd_ccf<- rd %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
ccf(rd_ccf$t7sum, rd_ccf$copies_mL) 
title("Gulf DustxCopies CCA", line =1)
```

**A negative Lag value indicates that x is a predictor of y, or in this case, dust is a predictor of copies_mL- so we will be looking at these negative values. A positive ACF indicates a positive relationship between the two variables, where a negative ACF indicates a negative relationship. According to the ccf plots, BO has a significant lag relationship a value of -1. In other words, dust occurs, and one day later, we see a response in copies_mL. C2 has a lag at -2, and RD at -1**

## Blind Oso Site:

#### Add one day lag

```{r}
bo2<- bo%>% mutate(
  lag =Lag(bo$t7sum, shift  = 1)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

#### Create a linear model

```{r}
#Create Recipe for Growth 
growth_recipe_bo<- recipe(copies_mL ~ lag, data = bo2)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe_bo)

#Use workflow to fit model to  data set
growth_fit_bo<- Growth_wflow %>%
  fit(data = bo2)

#View as Tibble 
growth_fit_bo %>%
  extract_fit_parsnip() %>%
  tidy()
```

### Use model to test data

```{r}
aug_test_bo <- augment(growth_fit_bo, bo2)
```

### RMSE and R2 to test model performance

```{r}
rmse <- aug_test_bo %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_bo %>% rsq(truth = copies_mL, .pred)
rd_metrics<- full_join(rmse, rsq)
```

### AIC to assess Performance

Approaches based on such selection criteria essentially try to guess how the model would perform if it were to be fit to new data, without actually trying to do it (in contrast to CV).

```{r}
bo_mod<- lm(copies_mL ~ lag, data = bo2)
AIC(bo_mod)
```

### Residuals

```{r}
res_bo<- resid(bo_mod)
plot(fitted(bo_mod), res_bo)
abline(0,0)
```

### Plot

```{r}
aug_test_bo %>% ggplot(aes(copies_mL, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (BO+1)")
```

## Canals:

#### Add two day lag

```{r}
c22<- c2%>% mutate(
  lag =Lag(c2$t7sum, shift  = 2)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

#### Create a linear model

```{r}
#Create Recipe for Growth 
growth_recipe_c2<- recipe(copies_mL ~ lag, data = c22)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe_c2)

#Use workflow to fit model to  data set
growth_fit_c2<- Growth_wflow %>%
  fit(data = c22)

#View as Tibble 
growth_fit_c2 %>%
  extract_fit_parsnip() %>%
  tidy()
```

### Use model to test data

```{r}
aug_test_c2 <- augment(growth_fit_c2, c22)
```

### RMSE and R2 to test model performance

```{r}
rmse <- aug_test_c2 %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_c2 %>% rsq(truth = copies_mL, .pred)
rd_metrics<- full_join(rmse, rsq)
```

### AIC to assess Performance

```{r}
c2_mod<- lm(copies_mL ~ lag, data = c22)
AIC(c2_mod)
```

### Residuals

```{r}
res_c2<- resid(c2_mod)
plot(fitted(c2_mod), res_c2)
abline(0,0)
```

### Plot

```{r}
aug_test_c2 %>% ggplot(aes(copies_mL, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (C2+2)")
```

## Gulf Site:

#### Add two day lag

```{r}
rd2<- rd%>% mutate(
  lag =Lag(rd$t7sum, shift  = 2)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

#### Create a linear model

```{r}
#Create Recipe for Growth 
growth_recipe_rd<- recipe(copies_mL ~ lag, data = rd2)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe_rd)

#Use workflow to fit model to  data set
growth_fit_rd<- Growth_wflow %>%
  fit(data = rd2)

#View as Tibble 
growth_fit_rd %>%
  extract_fit_parsnip() %>%
  tidy()
```

### Use model to test data

```{r}
aug_test_rd <- augment(growth_fit_rd, rd2)
```

### RMSE and R2 to test model performance

```{r}
rmse <- aug_test_rd %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_rd %>% rsq(truth = copies_mL, .pred)
rd_metrics<- full_join(rmse, rsq)
```

### AIC to assess Performance

```{r}
rd_mod<- lm(copies_mL ~ lag, data = rd2)
AIC(rd_mod)
```

### Residuals

```{r}
res_rd<- resid(rd_mod)
plot(fitted(rd_mod), res_rd)
abline(0,0)
```

### Plot

```{r}
aug_test_rd %>% ggplot(aes(lag, copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (RD+2)")
```

# Multivariate Models

[This code is still in progress. For now, please ignore:]{.ul}

#### Make New Dataset

```{r}
master2<- dust0 %>% full_join(data) %>%
  filter(hr_cst %in% "01") %>%
  select(site, sal, temp, amm, nn,orthop, copies_mL, t7sum) %>%
  na.omit()
```

#### Site-specific Data sets

```{r}
bo_a<- master2 %>%
  filter(site %in% "Blind Oso") %>%
  select(!site)

c2_a<- master2 %>%
  filter(site %in% "Canals")%>%
  select(!site)

rd_a<- master2 %>%
  filter(site %in% "Gulf")%>%
  select(!site)
```

## BO: Using all Variables vs. Growth

```{r}
#Create Recipe for Growth and all other parameters
growth_recipe_bo<- recipe(copies_mL ~ ., data = bo_a)

#Set up linear model
lm_mod_all<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod_all) %>%
  add_recipe(growth_recipe_bo)

#Use workflow to fit model to  data set
growth_fit_bo<- Growth_wflow %>%
  fit(data = bo_a)

#Pull model fit using parsnip()
growth_fit_bo %>%
  extract_fit_parsnip() %>%
  tidy()
```

Plot

```{r}
tidy(growth_fit_bo) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```
### Use model to test data

```{r}
aug_test_bo_a <- augment(growth_fit_bo, bo_a)
```

### RMSE and R2 to test model performance

```{r}
rmse <- aug_test_bo_a %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_bo_a %>% rsq(truth = copies_mL, .pred)
bo_a_metrics<- full_join(rmse, rsq)
```

### AIC to assess Performance

```{r}
bo_a_mod<- lm(copies_mL ~ ., data = bo_a)
AIC(bo_a_mod)
```

### Residuals

```{r}
res_bo_a<- resid(bo_a_mod)
plot(fitted(bo_a_mod), res_bo_a)
abline(0,0)
```
## C2: Using all Variables vs. Growth

```{r}
#Create Recipe for Growth and all other parameters
growth_recipe_c2<- recipe(copies_mL ~ ., data = c2_a)

#Set up linear model
lm_mod_all<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod_all) %>%
  add_recipe(growth_recipe_c2)

#Use workflow to fit model to  data set
growth_fit_c2<- Growth_wflow %>%
  fit(data = c2_a)

#View as Tibble 
growth_fit_c2 %>%
  extract_fit_parsnip() %>%
  tidy()
```

Plot

```{r}
tidy(growth_fit_c2) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```
### Use model to test data

```{r}
aug_test_c2_a <- augment(growth_fit_c2, c2_a)
```

### RMSE and R2 to test model performance

```{r}
rmse <- aug_test_c2_a %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_c2_a %>% rsq(truth = copies_mL, .pred)
c2_a_metrics<- full_join(rmse, rsq)
```

### AIC to assess Performance

```{r}
c2_a_mod<- lm(copies_mL ~ ., data = c2_a)
AIC(c2_a_mod)
```

### Residuals

```{r}
res_c2_a<- resid(c2_a_mod)
plot(fitted(c2_a_mod), res_c2_a)
abline(0,0)
```
## RD: Using all Variables vs. Growth

```{r}
#Create Recipe for Growth and all other parameters
growth_recipe_rd<- recipe(copies_mL ~ ., data = rd_a)

#Set up linear model
lm_mod_all<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod_all) %>%
  add_recipe(growth_recipe_rd)

#Use workflow to fit model to  data set
growth_fit_rd<- Growth_wflow %>%
  fit(data = rd_a)

#View as Tibble 
growth_fit_rd %>%
  extract_fit_parsnip() %>%
  tidy()
```

Plot

```{r}
tidy(growth_fit_rd) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```
### Use model to test data

```{r}
aug_test_rd_a <- augment(growth_fit_rd, rd_a)
```

### RMSE and R2 to test model performance

```{r}
rmse <- aug_test_rd_a %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_rd_a %>% rsq(truth = copies_mL, .pred)
rd_a_metrics<- full_join(rmse, rsq)
```

### AIC to assess Performance

```{r}
rd_a_mod<- lm(copies_mL ~ ., data = rd_a)
AIC(rd_a_mod)
```

### Residuals

```{r}
res_rd_a<- resid(rd_a_mod)
plot(fitted(rd_a_mod), res_rd_a)
abline(0,0)
```