### 2023-04-03 code/analysis_code/analysis_README.Rmd
Added LM for Nutrient Data

This folder contains code for analyzing cleaned data sets and aids in visualization of parameters of interest.\
(1) The dust subfolder contains `daily_dust.qmd`, which looks at dust concentration over the daily time series. It also contains `dust_and_copies.qmd` which combine Vibrio enumeration with dust data.\
(2) The env sub-folder contains `env_par.qmd` that examines temperature and salinity over the time series.\
(3) The qPCR sub-folder contains `daily_qmd` that has data on Vibrio enumeration. This is also seen in the `dust_and_copies.qmd` in the dust subfolder \
(4) The stats folder contains:
- `1_distributiom.qmd`: Checks distribution of variables of interest
- `2_cca.qmd`: Performs cross-correlation analysis to test for lags in time series (between dust input and Vibrio growth)
- `3_uni_variate.qmd` : runs univariate models on variable and predictors of interest.
- `4_multi_variate.qmd`: runs multivariate models on variable of interest (copies_mL)