### 2023-04-03 code/analysis_code/stats/4_multi_variate.qmd

---
title: "multi-variate"
format:
  html:
    theme: default
---

### Libraries

```{r, include = FALSE}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
library(dotwhisker) #Fpr visualization
library(lubridate)#Dates
```

# LOADING AND CLEANING DATA

### Data

```{r}
dust<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
data<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data
```

### Clean and Trim

```{r}
dust<- dust %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))

data<- data %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

### Combine into one big data set

```{r}
master<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.

master<- master %>%
  select(site, sal, temp, copies_mL,t7sum, date) #%>%
 # na.omit()
```

# Multivariate Models

[This code is still in progress. For now, please ignore:]{.ul}

### Site-specific Data sets and Clean df

```{r}
bo_a<- master %>%
  filter(site %in% "Blind Oso" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
bo_a<- bo_a %>%
mutate(
  lag =Lag(bo_a$t7sum, shift  = 1)) %>% #Create 1 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) %>%
select(!c(date,t7sum,site))

c2_a<- master %>%
  filter(site %in% "Canals" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
c2_a<- c2_a %>%
mutate(
  lag =Lag(c2_a$t7sum, shift  = 2)) %>% #Make 2 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))%>%
select(!c(date,t7sum,site))

rd_a<- master %>%
  filter(site %in% "Gulf" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06')) 
rd_a<- rd_a %>%
mutate(
  lag =Lag(rd_a$t7sum, shift  = 2)) %>% #Make 2 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))%>%
select(!c(date,t7sum,site))
```

## BO: Growth\~ Sal+Dust

```{r}
#Create Recipe for Growth and all other parameters
growth_recipe_bo<- recipe(copies_mL ~ sal+lag, data = bo_a)

#Set up linear model
lm_mod_all<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod_all) %>%
  add_recipe(growth_recipe_bo)

#Use workflow to fit model to  data set
growth_fit_bo<- Growth_wflow %>%
  fit(data = bo_a)

#Pull model fit using parsnip()
growth_fit_bo %>%
  extract_fit_parsnip() %>%
  tidy()
```

Plot

```{r}
tidy(growth_fit_bo) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```

### Use model to test data

```{r}
aug_test_bo_a <- augment(growth_fit_bo, bo_a)
```

### RMSE and R2 to test model performance

```{r}
rmse <- aug_test_bo_a %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_bo_a %>% rsq(truth = copies_mL, .pred)
bo_a_metrics<- full_join(rmse, rsq)
```

### AIC to assess Performance

```{r}
bo_a_mod<- lm(copies_mL ~ temp+sal+lag, data = bo_a)
AIC(bo_a_mod)
```

### Residuals

```{r}
res_bo_a<- resid(bo_a_mod)
plot(fitted(bo_a_mod), res_bo_a)
abline(0,0)
```

## C2: Using all Variables vs. Growth

```{r}
#Create Recipe for Growth and all other parameters
growth_recipe_c2<- recipe(copies_mL ~ ., data = c2_a)

#Set up linear model
lm_mod_all<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod_all) %>%
  add_recipe(growth_recipe_c2)

#Use workflow to fit model to  data set
growth_fit_c2<- Growth_wflow %>%
  fit(data = c2_a)

#View as Tibble 
growth_fit_c2 %>%
  extract_fit_parsnip() %>%
  tidy()
```

Plot

```{r}
tidy(growth_fit_c2) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```

### Use model to test data

```{r}
aug_test_c2_a <- augment(growth_fit_c2, c2_a)
```

### RMSE and R2 to test model performance

```{r}
rmse <- aug_test_c2_a %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_c2_a %>% rsq(truth = copies_mL, .pred)
c2_a_metrics<- full_join(rmse, rsq)
```

### AIC to assess Performance

```{r}
c2_a_mod<- lm(copies_mL ~ ., data = c2_a)
AIC(c2_a_mod)
```

### Residuals

```{r}
res_c2_a<- resid(c2_a_mod)
plot(fitted(c2_a_mod), res_c2_a)
abline(0,0)
```

## RD: Using all Variables vs. Growth

```{r}
#Create Recipe for Growth and all other parameters
growth_recipe_rd<- recipe(copies_mL ~ ., data = rd_a)

#Set up linear model
lm_mod_all<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod_all) %>%
  add_recipe(growth_recipe_rd)

#Use workflow to fit model to  data set
growth_fit_rd<- Growth_wflow %>%
  fit(data = rd_a)

#View as Tibble 
growth_fit_rd %>%
  extract_fit_parsnip() %>%
  tidy()
```

Plot

```{r}
tidy(growth_fit_rd) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```

### Use model to test data

```{r}
aug_test_rd_a <- augment(growth_fit_rd, rd_a)
```

### RMSE and R2 to test model performance

```{r}
rmse <- aug_test_rd_a %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_rd_a %>% rsq(truth = copies_mL, .pred)
rd_a_metrics<- full_join(rmse, rsq)
```

### AIC to assess Performance

```{r}
rd_a_mod<- lm(copies_mL ~ ., data = rd_a)
AIC(rd_a_mod)
```

### Residuals

```{r}
res_rd_a<- resid(rd_a_mod)
plot(fitted(rd_a_mod), res_rd_a)
abline(0,0)
```