### 2023-04-04 code/analysis_code/dust/daily_dusts.qmd
Clean scripts

---
title: "New Dust and Copies"
format:
  html:
    theme: default
---

**This code is useful for looking at variations of daily dust trends. For example, when running statistical analysis, it may be best to look at dust concentration at a specific hour, or the sum per day, or the average. We will be working with a sum that is 24hr prior to sample collection (t7sum). This utilizes a data set that has the variations of dust trends.**

# Getting Started

#### Libraries

```{r}
library(here)
library(tidyverse)
```

#### Data

```{r}
dust<- read_csv(here("data", "processed_data", "dust_master.csv")) 
```

# Cleaning

There are a few columns that we do not need or that contain NaN. So let's remove those.

#### Remove NAs

```{r}
dust<- dust %>%
  filter(!date %in% as.Date('2022-07-04'))
```

#### Select Dates for Daily Time series

```{r}
dust<-
  dust %>%
  filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19'))
```

### Plot T7sum over Daily Time series

<!--# Make this graph look nicer -->

```{r}
dust %>% ggplot()+ geom_line(
  aes(x = date,
      y = t7sum))+
  labs(title = "Dust Concentration over Daily Time Series",
       x = "Date",
       y = "Dust Concentration (AOD)")+
  theme_bw()
```