### 2023-04-04 code/wrangling_code/1_dust.qmd

---
title: "Untitled"
format:
  html:
    theme: default
---

**This code takes dust concentration data from NRL and cleans it. Steps include, switching to the correct timezone, creating concentration sums per day, as well as combining the dust data set with the vibrio enumeration data set.**

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(lubridate) #To work with dates 
library(here)
library(base) #For ColSum
```

#### Data

```{r}
nrl<- read_csv(here("data","raw_data","dust.csv")) #Dust concentration data set
copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) #Vibrio counts data set from qPCR

```

# Cleaning

Dust AOD was captured at four time points per day. We need to make a column that depicts that date AND hour of collection.

#### Make Year-Month-Day-Hour column from yyymmddtt column

```{r}
nrl1<- nrl %>% 
  mutate(
    ymdh = ymd_h(nrl$yyymmddtt))
```

This dust data set is in Coordinated Universal Time (UTC) and needs to be converted to Central Standard/Daylight Time (which is UTC-6 hours and UTC-5 hours for daylight savings.

#### Switch from UTC --> CST

```{r}
nrl2<- nrl1 %>%
  mutate(
    cst = as_datetime(nrl1$ymdh, tz= c("America/Chicago"))) #CST/CDT
```

#### Make column with just date in CST

```{r}
#Just the data from CST/CDT
nrl3<- nrl2 %>%
  mutate(
    just_date= as.Date(substr(nrl2$cst, 1,10)))

#Create a column for hr in cst
dust_ind<- nrl3 %>%
  mutate(hr_cst = substr(nrl3$cst, 12,13)) %>%
  select(hr_cst, just_date, dust) %>% #Make final DF with dust, hours, and time in CST
  rename("date" = "just_date")
```

Now `dust_ind` data set has the date and hour of collection in CST/CDT, and the concentrations (AOD) at each time point\

For future analyses, it may be helpful to have the dust sum of each day (T0 +T6 +T12 +T18). This could give us an estimate for the dust concentration in a single day.

#### Dust Sum

```{r}
dust_sum<- aggregate(dust_ind$dust, by=list(date=dust_ind$date),FUN=sum) %>% #could also use groupby
 rename("SUM" = "x")
```

This data set `dust_sum` has the dust concentration summed for each calendar day. Now let's combine the dust SUM to the dust INDIVUDUAL data frames\

#### Combine Sum and Individual Data frames

```{r}
dust_master<- dust_sum %>% right_join(dust_ind)
```

#### Save as RDS

```{r}
saveRDS(dust_master, file = here("data", "processed_data", "dust_clean.rds")) #Dust with Cst Time and Hours

```