### 2023-04-04 code/analysis_code/env/env_par.qmd
cleaning up

---
title: "Environmental Parameters"
format:
  html:
    theme: default
---

**This data set contains environmental parameters: Temp and Salinity, among other environmental parameters.**

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(here)
library(ggpubr)
library(lubridate)
```

#### Load Data

```{r}
d1<- read_csv(here("data", "processed_data", "tx_master.csv"))
```

# Cleaning

#### Daily data frame

```{r}
daily<- d1 %>%
  filter(date %in% c(as.Date("2022-07-07"):as.Date("2022-07-19"))) %>%
  select(!...1)
```

# Visualization

#### Plotting Daily Temperature

```{r}
temp<- daily %>% ggplot() + geom_line(
  aes(
    x = date,
    y = temp,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Temperature (°C)",
    title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 15),
    legend.position = c(0.1,0.2),
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

temp

```

#### Plotting Daily Salinity

```{r}
sal<- daily %>% ggplot() + geom_line(
  aes(
    x = date,
    y = sal,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Salinity",
    title = "Salinity across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    legend.position = "none")+
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

sal
```

#### Combine Temperature and Salinity Plots

```{r, fig.width= 4, fig.height=4}
fig<- ggarrange(temp,sal + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")

fig
```

#### Save as PNG

```{r}
png(file = here("results", "plots", "temp_sal.png"))
fig
dev.off()
```