### 2023-04-04 code/analysis_code/qPCR/daily.qmd

---
title: "Exploratory"
format:
  html:
    theme: default
---

**This code looks at Vibrio enumeration over the daily time series. This information can also be found in `dust_and_copies.qmd`**

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(here)
```

#### Data

```{r}
data<- read_csv(here("data", "processed_data", "tx_master.csv")) 
```

# Cleaning

#### Make mini dataframes for each site

```{r}
#Daily All Sites
daily<- data %>%
  select(date,copies_mL,site,Type) %>%
  filter(Type %in% "Daily") %>%
drop_na() #Get rid of NAs

#Daily by Site
bo_daily <- daily %>% filter(site %in% "Blind Oso") #Blind Oso Bay
c2_daily <- daily %>% filter(site %in% "Canals") #Canals
rd_daily <- daily %>% filter(site %in% "Gulf") #Gulf Site

```

# Visualization

#### Total Daily Vibrio at Gulf

```{r}
rd_daily %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "darkblue") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at Gulf Site") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Total Daily Vibrio at Canals

```{r}
c2_daily %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "darkorange") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at Canals") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Total Daily Vibrio at Blind Oso

```{r}
bo_daily %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "darkgreen") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at Blind Oso") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Total Daily Vibrio

```{r}
tot_vib<- daily %>% ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site))+
  facet_wrap(~site) +
  theme_bw()+
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at all Sites") +
  theme(plot.title = element_text(hjust = 0.5)) + #center title
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + #Tilt x axis labels 
  theme(legend.position = "none") #Removes legend

tot_vib
```

#### Save as PNG

```{r}
png(file = here("results", "plots", "tot_vib_allsites_daily.png"))
tot_vib
dev.off()
```