### 2023-04-06 code/analysis_code/env/env_par.qmd
Addedto env data viz

---
title: "Environmental Parameters"
format:
  html:
    theme: default
---

**This data set contains environmental parameters: Temp and Salinity, among other environmental parameters.**

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(here)
library(ggpubr)
library(lubridate)
```

#### Load Data

```{r}
d1<- read_csv(here("data", "processed_data", "tx_master.csv"))
```

# Cleaning

#### Daily data frame

```{r}
daily<- d1 %>%
  filter(date %in% c(as.Date("2022-07-07"):as.Date("2022-07-19"))) %>%
  filter(!date %in% as.Date('2022-07-10')) %>%
  select(!...1)
```

# Visualization

#### Temperature

```{r}
temp<- daily %>% ggplot() + geom_line(
  aes(
    x = date,
    y = temp,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Temperature (°C)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

#### Salinity

```{r}
sal<- daily %>% ggplot() + geom_line(
  aes(
    x = date,
    y = sal,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Salinity",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

#### Dissolved Oxygen

```{r}
do<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = do_mgl,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "DO (mg/L)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

#### Nitrite:Nitrate

```{r}
nn<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = nn,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Nitrate-Nitrite (µM)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

#### Orthophosphate

```{r}
ortho<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = orthop,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Orthophosphate (µM)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    axis.title.x = element_text(size = 10),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

#### DON

```{r}
don<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = don,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "DON (µM)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    axis.title.x = element_text(size = 10),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

#### DOC

```{r}
doc<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = doc,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "DOC (µM)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    axis.title.x = element_text(size = 10),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

#### pH

```{r}
pH<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = pH,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "pH",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_text(size = 10),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

#### Ammonium

```{r}
amm<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = amm,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Ammonium (µM)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

#### Total Nitrogen

```{r}
tn<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = tn,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Total Nitrogen (µM)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

#### Silicate

```{r}
sil<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = sil,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Silicate (µM)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

#### Copies/mL

```{r}
cop<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = copies_mL,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Vibrio (copies per mL)",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.key.size = unit(0.3,'cm'),
    legend.position = c(0.2,0.7),
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid'),
    legend.text = element_text(size = 7)) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

#### Combine Environment Plots

```{r, fig.width= 10, fig.height= 5}
fig<- ggarrange(cop,temp,sal,sil,do,amm,nn,tn,ortho,doc, pH,don + font("x.text", size = 10), ncol = 4, nrow = 3,align = "v")

fig
```

# Save as PNG

```{r}
png(file = here("results", "plots", "temp_sal.png"))
fig
dev.off()
```