### 2023-04-06 code/analysis_code/stats/cor.qmd
Cor Plots

---
title: "cor"
format:
  html:
    theme: default
---

**This code creates a correlation matrix to look at relationships between variables.**

[**THIS CODE IS STILL IN THE WORKS**]{.ul}

Link: http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(here)
library(Hmisc)#Cor Tables
library(corrplot) #Cor Plots
```

#### Data

```{r, include = FALSE}
dust0<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust permutations
data0<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data containing copies_mL and nutrients 
```

# Cleaning

### Clean and Trim

```{r}
dust<- dust0 %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) %>%
  mutate(t7sum = log10(t7sum)) #Log transform Dust conc (as it is not normal)

data<- data0 %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

### Combine into one big data set

```{r}
master0<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") %>% #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.
  select(!c(date,hr_cst,...1,Type, t1,t7,t13,t19,tsum,tavg,SUM,t7sum,t7avg))

```

#### Site DF

```{r}
bo<- 
  master0 %>%
  filter(site %in% "Blind Oso") %>%
  select(!site)

c2<- 
  master0 %>%
  filter(site %in% "Canals") %>%
  select(!site)

rd<- 
  master0 %>%
  filter(site %in% "Gulf") %>%
  select(!site)
```

# Correlation Matrix

#### Compute Correlation Matrix

```{r}
b<- 
  cor(bo)
c<- 
  cor(c2)
r<- cor(rd)
```

#### Make Matrix

```{r}
bo2<- rcorr(as.matrix(bo))
c22<-

```

#### Function to Format

```{r}
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
```

#### Look at Matrix again

```{r}
b<- rcorr(as.matrix(bo))
flattenCorrMatrix(b$r, b$P)
```

#### Visualize

```{r}
corrplot(b, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

corrplot(c, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

corrplot(r, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```