### 2023-04-07 code/analysis_code/analysis_README.Rmd
Worked with R2 pearson and wrapped up Correlation plots

This folder contains code for analyzing cleaned data sets and aids in visualization of parameters of interest.\
(1) The dust subfolder contains `daily_dust.qmd`, which looks at dust concentration over the daily time series. It also contains `dust_and_copies.qmd` which combine Vibrio enumeration with dust data.\
(2) The env sub-folder contains `env_par.qmd` that examines temperature and salinity over the time series.\
(3) The qPCR sub-folder contains `daily_qmd` that has data on Vibrio enumeration. This is also seen in the `dust_and_copies.qmd` in the dust subfolder \
(4) The stats folder contains:
- `1_distributiom.qmd`: Checks distribution of variables of interest
- `2_cca.qmd`: Performs cross-correlation analysis to test for lags in time series (between dust input and Vibrio growth)
- `3_correlation.qmd`: Runs Correlation matrices on all variables
- `4_uni_variate.qmd` : runs univariate models on variable and predictors of interest.
- `5_multi_variate.qmd`: runs multivariate models on variable of interest (copies_mL)