### 2023-04-07 code/analysis_code/stats/3_correlation.qmd

---
title: "cor"
format:
  html:
    theme: default
---

**This code creates a correlation matrix to look at relationships between variables. This helps identify specific relationships of interest**

[**THIS CODE IS STILL IN THE WORKS**]{.ul}

[Link](http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software):

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(here)
library(Hmisc)#Cor Tables and Lag
library(corrplot) #Cor Plots
library(lubridate) #For dates
```

#### Data

```{r, include = FALSE}
dust0<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust permutations
data0<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data containing copies_mL and nutrients 
```

# Cleaning

### Trim data sets from 7/4 to 7/19

```{r}
dust<- dust0 %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) %>%
  mutate(t7sum = log10(t7sum)) #Log transform Dust conc (as it is not normally distributed)

data<- data0 %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

### Combine into one big data set

```{r}
master0<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") %>% #While this is technically not for hour 1, this gets ride of all   duplicates for the sake of analysis.
  select(!c(hr_cst,...1,Type, t1,t7,t13,t19,tsum,tavg,SUM,dust,t7avg)) %>%
  rename("dust" = "t7sum") #Now dust variable is the t7sum
```

#### Make Lag data sets by Site

This cleans out non-numeric values, creates site lag from CCA, and producea a final daily plot

```{r}
bo<- 
  master0 %>%
  filter(site %in% "Blind Oso"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06"))) %>%
  select(!site) %>% 
  mutate(dust =Lag(dust, shift  = 1)) %>% #Lag from cca.qmd
  filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19')) %>%
  select(!date)

c2<- 
  master0 %>%
  filter(site %in% "Canals"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06"))) %>%
  select(!site) %>% 
  mutate(dust =Lag(dust, shift  = 2)) %>%
  filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19')) %>%
  select(!date)

rd<- 
  master0 %>%
  filter(site %in% "Gulf"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06"))) %>%
  select(!site) %>% 
  mutate(dust =Lag(dust, shift  = 1)) %>%
  filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19')) %>%
  select(!date)

```

# Correlation Matrix

There are different methods for **correlation analysis** : **Pearson parametric correlation test**, **Spearman** and **Kendall** rank-based **correlation analysis**. The default is [pearson correlation coefficient](http://www.sthda.com/english/wiki/correlation-test-between-two-variables) which measures the linear **dependence** between two variables. [kendall and spearman](http://www.sthda.com/english/wiki/correlation-test-between-two-variables) correlation methods are non-parametric **rank-based correlation test**.

#### Compute Correlation Matrix

If your data contain missing values, use the following R code to handle missing values by case-wise deletion.

    cor(x, method = "pearson", use = "complete.obs")

```{r, include = FALSE}
b<- cor(bo)
round(b,2) #Round correlation coeff to "x" digits

c<- cor(c2)
round(c,2)

r<- cor(rd)
round(r,2)
```

#### Make Function to Format

Simple function for formatting a **correlation matrix** into a table with 4 columns containing : Column 1 : row names (variable 1 for the correlation test), Column 2 : column names (variable 2 for the correlation test), Column 3 : the **correlation coefficients**, Column 4 : the **p-values** of the correlations

```{r}
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut])}
```

#### Make Matrix

The function **rcorr()** \[in **Hmisc** package\] can be used to compute the **significance levels** for **pearson** and **spearman correlations**. It returns both the correlation coefficients and the p-value of the correlation for all possible pairs of columns in the data table.

```{r}
bclean<- rcorr(as.matrix(bo))
flattenCorrMatrix(bclean$r, bclean$P)

cclean<- rcorr(as.matrix(c2))
flattenCorrMatrix(cclean$r, cclean$P)

rclean<- rcorr(as.matrix(rd))
flattenCorrMatrix(rclean$r, rclean$P)
```

#### Visualize with `corrplot()` function

The function **corrplot()** takes the **correlation matrix** as the first argument. The second argument (type="upper") is used to display only the upper triangular of the **correlation matrix**.

```{r}
corrplot(b, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45,
         title = "Blind Oso",  mar=c(0,0,1,0)) #Adds and lowers title

corrplot(c, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45,
         title = "Canals", mar=c(0,0,1,0))

corrplot(r, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45,
         title = "Gulf", mar=c(0,0,1,0))
```

It's also possible to **combine correlogram with the significance test**. We'll use the result *res.cor2* generated in the previous section with **rcorr**() function \[in **Hmisc** package\]:

<!--# Get this to work -->

```{r}
# Insignificant correlation are crossed
#corrplot(bclean$r, type="upper", order="hclust", 
        # p.mat = bclean$P, sig.level = 0.01, insig = "blank")
# Insignificant correlations are leaved blank
#corrplot(bclean$r, type="upper", order="hclust", 
        # p.mat = bclean$P, sig.level = 0.01, insig = "blank")
```