### 2023-04-17 code/analysis_code/stats/5_multi_variate_models/rd_multi.qmd

---
title: "multi-variate"
format:
  pdf:
    theme: default
---

# Getting Started

#### Libraries

```{r, include = FALSE}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
library(dotwhisker) #Fpr visualization
library(lubridate)#Dates
library(kableExtra) #Tables
```

#### Data

```{r}
dust<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
data<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data
```

# Cleaning

#### Clean and Trim

```{r}
dust<- dust %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))

data<- data %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

#### Make into one data set

```{r}
master<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.

master<- master %>%
  select(site, sal, temp, copies_mL,t7sum, date, orthop,nn, don,doc) %>%
  mutate(t7sum = log10(t7sum))
```

#### Site-specific Data sets and Clean df

```{r}
b<- master %>%
  filter(site %in% "Gulf" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
b<- b %>%
mutate(
  lag =Lag(b$t7sum, shift  = 1)) %>% #Create 1 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) %>%
select(!c(date,t7sum,site)) %>%
  rename("log10dust" = "lag")
```

# Models

## 0. Null Model:

### Create 5 Fold

```{r}
fold <- vfold_cv(b, v = 5, repeats = 5) #Data was too small to stratify (<20)
```

### Make Models

```{r, include= FALSE}
#Create Null Recipe for Growth at all three sites
null<- recipe(copies_mL ~ 1, data = b)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model for three sites
null_wflow<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(null)

#Use workflow to fit model to  data set
null_fit<-
  fit_resamples(null_wflow, resamples = fold)

```

### Get Metrics for Null

#### RMSE

```{r}
null_met <- 
  collect_metrics(null_fit)
null_met
rm0<- 50771.79
rs0<- "NA"
```

#### AIC

```{r}
null_mod<- 
  lm(copies_mL ~ 1, data = b)
aic0<-AIC(null_mod) %>% round(digits = 2)
```

## 1. Dust Model

#### Model

```{r}
#Create Recipe for Growth 
growth_recipe1<- recipe(copies_mL ~ log10dust, data = b)

#Workflow that adds recipe to model
Growth_wflow1<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe1)

#Use workflow to fit model to  data set
growth_fit1<- Growth_wflow1 %>%
  fit(data = b)

#View as Tibble 
growth_fit1 %>%
  extract_fit_parsnip() %>%
  tidy()
```

#### Metrics

```{r}
aug_test1 <- augment(growth_fit1, b)
rmse1 <- aug_test1 %>% rmse(truth = copies_mL, .pred)
rsq1 <- aug_test1 %>% rsq(truth = copies_mL, .pred)

mod1<- lm(copies_mL ~ log10dust, b)
aic1<-AIC(mod1) %>% round(digits = 2)

metrics1<- full_join(rmse1, rsq1)
metrics1
```

## 2. Water Chemistry Model

#### Model

```{r}
#Create Recipe for Growth 
growth_recipe2<- recipe(copies_mL ~ sal+temp, data = b)

#Workflow that adds recipe to model
Growth_wflow2<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe2)

#Use workflow to fit model to  data set
growth_fit2<- Growth_wflow2 %>%
  fit(data = b)

#View as Tibble 
growth_fit2 %>%
  extract_fit_parsnip() %>%
  tidy()
```

#### Metrics

```{r}
aug_test2 <- augment(growth_fit2, b)
rmse2 <- aug_test2 %>% rmse(truth = copies_mL, .pred)
rsq2 <- aug_test2 %>% rsq(truth = copies_mL, .pred)

mod2<- lm(copies_mL ~ sal+temp, b)
aic2<- AIC(mod2)%>% round(digits = 2)

metrics2<- full_join(rmse2, rsq2)
metrics2
```

## 3. Chemistry + Dust

#### Model

```{r}
#Create Recipe for Growth 
growth_recipe3<- recipe(copies_mL ~ sal+temp + log10dust, data = b)

#Workflow that adds recipe to model
Growth_wflow3<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe3)

#Use workflow to fit model to  data set
growth_fit3<- Growth_wflow3 %>%
  fit(data = b)

#View as Tibble 
growth_fit3 %>%
  extract_fit_parsnip() %>%
  tidy()
```

#### Metrics

```{r}
aug_test3 <- augment(growth_fit3, b)
rmse3 <- aug_test3 %>% rmse(truth = copies_mL, .pred)
rsq3 <- aug_test3 %>% rsq(truth = copies_mL, .pred)

mod3<- lm(copies_mL ~ sal+temp + log10dust, b)
aic3<- AIC(mod3)%>% round(digits = 2)

metrics3<- full_join(rmse3, rsq3)
metrics3
```

## 4. Model 3 + Nutrients

#### Model

```{r}
#Create Recipe for Growth 
growth_recipe4<- recipe(copies_mL ~ sal+temp+orthop+log10dust+nn, data = b)

#Workflow that adds recipe to model
Growth_wflow4<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe4)

#Use workflow to fit model to  data set
growth_fit4<- Growth_wflow4 %>%
  fit(data = b)

#View as Tibble 
growth_fit4 %>%
  extract_fit_parsnip() %>%
  tidy()
```

#### Metrics

```{r}
aug_test4 <- augment(growth_fit4, b)
rmse4 <- aug_test4 %>% rmse(truth = copies_mL, .pred)
rsq4 <- aug_test4 %>% rsq(truth = copies_mL, .pred)

mod4<- lm(copies_mL ~ sal+temp+orthop+log10dust+nn, b)
aic4<- AIC(mod4)%>% round(digits = 2)

metrics4<- full_join(rmse4, rsq4)
metrics4
```

## 5. Module 4 + DOM

#### Model

```{r}
#Create Recipe for Growth 
growth_recipe5<- recipe(copies_mL ~ sal+temp+log10dust+orthop+nn+don+doc, data = b)

#Workflow that adds recipe to model
Growth_wflow5<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe5)

#Use workflow to fit model to  data set
growth_fit5<- Growth_wflow5 %>%
  fit(data = b)

#View as Tibble 
growth_fit5 %>%
  extract_fit_parsnip() %>%
  tidy()
```

#### Metrics

```{r}
aug_test5 <- augment(growth_fit5, b)
rmse5 <- aug_test5 %>% rmse(truth = copies_mL, .pred)
rsq5 <- aug_test5 %>% rsq(truth = copies_mL, .pred)

mod5<- lm(copies_mL ~ sal+temp+log10dust+orthop+nn+don+doc, b)
aic5<-AIC(mod5)%>% round(digits = 2)

metrics5<- full_join(rmse5, rsq5)
metrics5
```

# Tables

#### Label Metrics

```{r}
#R2
rs1<- rsq1$.estimate %>% round(digits = 2)
rs2<- rsq2$.estimate %>% round(digits = 2)
rs3<- rsq3$.estimate %>% round(digits = 2)
rs4<- rsq4$.estimate %>% round(digits = 2)
rs5<- rsq5$.estimate %>% round(digits = 2)

#RMSE
rm1<- rmse1$.estimate 
rm2<- rmse2$.estimate
rm3<- rmse3$.estimate
rm4<- rmse4$.estimate
rm5<- rmse5$.estimate
```

#### Make Table

```{r}
kable(data.frame(Model = c("Null Model", "Model 1: Dust", "Model 2: Water Chemistry", "Model 3: Chemistry + Dust", "Model 4: Model 3 + Nutrients", "Model 5: Model 4 + DOM"),
             R2 = c(rs0,rs1,rs2,rs3,rs4,rs5),
             RMSE = c(rm0,rm1,rm2,rm3,rm4,rm5),
             AIC = c(aic0, aic1,aic2,aic3,aic4,aic5)),
      caption = "Gulf: Comparative linear regression analysis of Vibrio copies per mL between 6 operational models.")
```

####