### 2023-04-19 code/analysis_code/stats/5_multi_variate_models/rf_models.qmd
+ RF model

---
title: "lasso"
format:
  html:
    theme: default
---

# Getting Started

#### Libraries

```{r, include = FALSE}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
library(dotwhisker) #Fpr visualization
library(lubridate)#Dates
library(kableExtra) #Tables
library(glmnet)
```

#### Data

```{r}
dust<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
data<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data
```

# Cleaning

#### Clean and Trim

```{r}
dust<- dust %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))

data<- data %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

#### Make into one data set

```{r}
master<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.

master<- master %>%
  select(site, sal, temp, copies_mL,t7sum, date, orthop,nn, don,doc) %>%
  mutate(t7sum = log10(t7sum))
```

#### Site-specific Data sets and Clean df

```{r}
b<- master %>%
  filter(site %in% "Blind Oso" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
b<- b %>%
mutate(
  lag =Lag(b$t7sum, shift  = 1)) %>% #Create 1 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) %>%
select(!c(date,t7sum,site)) %>%
  rename("log10dust" = "lag")

c<- master %>%
  filter(site %in% "Canals" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
c<- c %>%
mutate(
  lag =Lag(c$t7sum, shift  = 2)) %>% #Create 2 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) %>%
select(!c(date,t7sum,site)) %>%
  rename("log10dust" = "lag")

r<- master %>%
  filter(site %in% "Gulf" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
r<- r %>%
mutate(
  lag =Lag(r$t7sum, shift  = 1)) %>% #Create 1 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) %>%
select(!c(date,t7sum,site)) %>%
  rename("log10dust" = "lag")
```

# Models

## 0. Null Model:

### Create 5 Fold

```{r}
fold_bo <- vfold_cv(b, v = 5, repeats = 5) 
fold_c2 <- vfold_cv(c, v = 5, repeats = 5) 
fold_rd <- vfold_cv(r, v = 5, repeats = 5) 
```

### Make Models

```{r, include= FALSE}
#Create Null Recipe for Growth at all three sites
null_bo<- recipe(copies_mL ~ 1, data = b) 
null_c2<- recipe(copies_mL ~ 1, data = c) 
null_rd<- recipe(copies_mL ~ 1, data = r)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model for three sites
null_wflow_bo<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(null_bo)

null_wflow_c2<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(null_c2)

null_wflow_rd<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(null_rd)

#Use workflow to fit model to  data set
null_fit_bo<-
  fit_resamples(null_wflow_bo, resamples = fold_bo)

null_fit_c2<-
  fit_resamples(null_wflow_c2, resamples = fold_c2)

null_fit_rd<-
  fit_resamples(null_wflow_rd, resamples = fold_rd)

```

### Get Metrics for Null

#### RMSE

```{r}
null_met_bo <- 
  collect_metrics(null_fit_bo)
null_met_bo

null_met_c2 <- 
  collect_metrics(null_fit_c2)
null_met_c2

null_met_rd <- 
  collect_metrics(null_fit_rd)
null_met_rd
```

#### AIC

```{r}
bo_null_mod<- 
  lm(copies_mL ~ 1, data = b)
AIC(bo_null_mod) %>% round(digits = 2)

c2_null_mod<- 
  lm(copies_mL ~ 1, data = c)
AIC(c2_null_mod) %>% round(digits = 2)

rd_null_mod<- 
  lm(copies_mL ~ 1, data = r)
AIC(rd_null_mod) %>% round(digits = 2)
```

## Blind Oso Random Forest

#### Recipe

```{r}
rec<- recipe(copies_mL ~ ., data = b)
```

```{r}
cores <- parallel::detectCores()
cores
```

#### Specify Model

```{r}
rf_mod <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
  set_engine("ranger", num.threads = cores) %>% 
  set_mode("regression")
```

#### Creating Workflow

```{r}
rf_wf <- workflow() %>%
  add_model(rf_mod) %>%
  add_recipe(rec)
```

#### Create Tuning Grid

```{r}
rf_grid  <- expand.grid(mtry = c(3, 4, 5, 6),
                        min_n = c(40,50,60), 
                        trees = c(500,1000)  )
```

#### Cross-validation

```{r}
rf_resample <- 
  rf_wf %>% 
  tune_grid(fold_bo,
            grid = 25,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(rmse))
```

```{r}
rf_resample %>%
  collect_metrics()
```

##### Plot Model Performance

```{r}
#Plot of actual train data
rf_resample %>%
  autoplot()
```

##### Showing and Selecting Best Performing Models

```{r}
#Showing best performing tree models
rf_resample %>%
  show_best(n=1)

#Selects best performing model
best_rf <- rf_resample %>%
  select_best(method = "rmse")
```

##### Create Final Fit

```{r}
rf_final_wf <- 
  rf_wf %>% 
  finalize_workflow(best_rf)

#Create workflow for fitting model to train_data2 predictions
rf_final_fit <- 
  rf_final_wf %>%
  fit(b) 
```

##### Calculate Residuals

```{r}
rf_residuals <- rf_final_fit %>%
  augment(b) %>% #use augment() to make predictions from train data
  select(c(.pred, copies_mL)) %>%
  mutate(.resid = copies_mL - .pred) #calculate residuals and make new row.

rf_residuals
```

##### Model Predictions from Tuned Model vs Actual Outcomes

```{r}
rf_pred_plot <- ggplot(rf_residuals, 
                          aes(x = copies_mL, 
                              y = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Actual: Random Forest", 
       x = "Copies/mL Actual", 
       y = "Copies/mL Prediction")
rf_pred_plot

rf_residual_plot <- ggplot(rf_residuals, 
                              aes(y = .resid, 
                                  x = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Residuals: Random Forest", 
       x = "Copies/mL Prediction", 
       y = "Residuals")
plot(rf_residual_plot) #view plot
```

## Canals Random Forest

#### Recipe

```{r}
rec<- recipe(copies_mL ~ ., data = c)
```

```{r}
cores <- parallel::detectCores()
cores
```

#### Specify Model

```{r}
rf_mod <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
  set_engine("ranger", num.threads = cores) %>% 
  set_mode("regression")
```

#### Creating Workflow

```{r}
rf_wf <- workflow() %>%
  add_model(rf_mod) %>%
  add_recipe(rec)
```

#### Create Tuning Grid

```{r}
rf_grid  <- expand.grid(mtry = c(3, 4, 5, 6),
                        min_n = c(40,50,60), 
                        trees = c(500,1000)  )
```

#### Cross-validation

```{r}
rf_resample <- 
  rf_wf %>% 
  tune_grid(fold_c2,
            grid = 25,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(rmse))
```

```{r}
rf_resample %>%
  collect_metrics()
```

##### Plot Model Performance

```{r}
#Plot of actual train data
rf_resample %>%
  autoplot()
```

##### Showing and Selecting Best Performing Models

```{r}
#Showing best performing tree models
rf_resample %>%
  show_best(n=1)

#Selects best performing model
best_rf <- rf_resample %>%
  select_best(method = "rmse")
```

##### Create Final Fit

```{r}
rf_final_wf <- 
  rf_wf %>% 
  finalize_workflow(best_rf)

#Create workflow for fitting model to train_data2 predictions
rf_final_fit <- 
  rf_final_wf %>%
  fit(c) 
```

##### Calculate Residuals

```{r}
rf_residuals <- rf_final_fit %>%
  augment(c) %>% #use augment() to make predictions from train data
  select(c(.pred, copies_mL)) %>%
  mutate(.resid = copies_mL - .pred) #calculate residuals and make new row.

rf_residuals
```

##### Model Predictions from Tuned Model vs Actual Outcomes

```{r}
rf_pred_plot <- ggplot(rf_residuals, 
                          aes(x = copies_mL, 
                              y = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Actual: Random Forest", 
       x = "Copies/mL Actual", 
       y = "Copies/mL Prediction")
rf_pred_plot

rf_residual_plot <- ggplot(rf_residuals, 
                              aes(y = .resid, 
                                  x = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Residuals: Random Forest", 
       x = "Copies/mL Prediction", 
       y = "Residuals")
plot(rf_residual_plot) #view plot
```

## Gulf Random Forest

#### Recipe

```{r}
rec<- recipe(copies_mL ~ ., data = r)
```

```{r}
cores <- parallel::detectCores()
cores
```

#### Specify Model

```{r}
rf_mod <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
  set_engine("ranger", num.threads = cores) %>% 
  set_mode("regression")
```

#### Creating Workflow

```{r}
rf_wf <- workflow() %>%
  add_model(rf_mod) %>%
  add_recipe(rec)
```

#### Create Tuning Grid

```{r}
rf_grid  <- expand.grid(mtry = c(3, 4, 5, 6),
                        min_n = c(40,50,60), 
                        trees = c(500,1000)  )
```

#### Cross-validation

```{r}
rf_resample <- 
  rf_wf %>% 
  tune_grid(fold_rd,
            grid = 25,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(rmse))
```

```{r}
rf_resample %>%
  collect_metrics()
```

##### Plot Model Performance

```{r}
#Plot of actual train data
rf_resample %>%
  autoplot()
```

##### Showing and Selecting Best Performing Models

```{r}
#Showing best performing tree models
rf_resample %>%
  show_best(n=1)

#Selects best performing model
best_rf <- rf_resample %>%
  select_best(method = "rmse")
```

##### Create Final Fit

```{r}
rf_final_wf <- 
  rf_wf %>% 
  finalize_workflow(best_rf)

#Create workflow for fitting model to train_data2 predictions
rf_final_fit <- 
  rf_final_wf %>%
  fit(r) 
```

##### Calculate Residuals

```{r}
rf_residuals <- rf_final_fit %>%
  augment(r) %>% #use augment() to make predictions from train data
  select(c(.pred, copies_mL)) %>%
  mutate(.resid = copies_mL - .pred) #calculate residuals and make new row.

rf_residuals
```

##### Model Predictions from Tuned Model vs Actual Outcomes

```{r}
rf_pred_plot <- ggplot(rf_residuals, 
                          aes(x = copies_mL, 
                              y = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Actual: Random Forest", 
       x = "Copies/mL Actual", 
       y = "Copies/mL Prediction")
rf_pred_plot

rf_residual_plot <- ggplot(rf_residuals, 
                              aes(y = .resid, 
                                  x = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Residuals: Random Forest", 
       x = "Copies/mL Prediction", 
       y = "Residuals")
plot(rf_residual_plot) #view plot
```