### 2023-05-01 data/processed_data/processed_data_README.Rmd

------------------------------------------------------------------------

This folder contains processed data that has been cleaned from code in the `wrangling` folder.\

## `total_clean_copies.csv`

Vibrio enumeration as gene copies per mL across time series

#### Variables

date, site, type (monthly or daily time series), and copies_mL.

## `dust_clean.rds`

Cleaned dust data

#### Variables

date, SUM (AOT summed by date), hr_cst (hour of collection in CST), and dust (AOT)

## `ysi_2.csv`

Environmental parameter data-- did not need cleaning

#### Variables

site, date, chlorophyll (chl), salinity (sal), temperature (temp), Dissolved Oxygen in mg/L (do_mgl), Dissolved Oxygen in % (do_per), pH, secchi depth (secchi), ammonium (amm), Nitrite+Nitrate (nn), orthophosphate (orthop), silicate (sil), dissolved inorganic nitrogen (din), dissolved inorganic nitrogen:dissolved inorganic phosphorous (din_dip), total dissolved nitrogen (tdn), dissolved organic nitrogen (don), dissolved organic carbon (doc), total organic carbon (toc), total nitrogen (tn).

## `tx_master.csv`

A compilation of all cleaned variables mentioned above

## `dust_master.csv`

Manually created dust data examining different dust permutations: time points, sums, averages (helpful for statistics)

#### Variables

date, time at collection 1 hour (t1), 7 hour (t7), 13 hour (t13), and 19 hour (t19). Sum of all 4 time points (tsum), average of all four time points (tavg), 24 hour hour composite sum (t7sum), 24 hour composite average (t7sum).