### 2023-05-01 data/raw_data/rawdata_README.Rmd

This folder contains two raw data sets:

## `dust.csv`

#### Variables:

yyymmddtt: Year_month_day_time of collection

dust: dust aerosol optical density. Concentrations (AOT) were collected at 4 time points (UTC): T0,T6,T12, and T18.

smoke: smoke

seasalt: seasalt

SS Total AOT: Total aerosol optical thickness

date: date

type: denotes daily sampling time period

## `copies_master.csv`

#### Variables:

Type: Denoted daily sampling period

Sample: Site and Day short code. Ex. BO11 = Blind Oso on 7/11/23

Cq: Cycle threshold value from qPCR. Indicated the cycle number that amplification was detected. A lower Cq value indicates earlier detection, and a higher concentration of target,\