### 2023-05-06 code/analysis_code/dust/daily_dusts.qmd
Final Edits
After double checking manuscript, I noticed that figure proportions were off. (small text etc). I went back and readjusted but am wondering why I am just now noticing this. I am working on my laptop (as opposed to my desktop), so i am wondering if R is somehow correcting for screen resolution, and I am seeing different sized plots due to this. Please let me know if my figures look off when you review, thanks!

---
title: "New Dust and Copies"
format:
  html:
    theme: default
---

**This code is useful for looking at variations of daily dust trends. For example, when running statistical analysis, it may be best to look at dust concentration at a specific hour, or the sum per day, or the average. We will be working with a sum that is 24hr prior to sample collection (t7sum). This utilizes a data set that has the variations of dust trends.**

# Getting Started

#### Libraries

```{r}
library(here)
library(tidyverse)
```

#### Data

```{r}
dust<- read_csv(here("data", "processed_data", "dust_master.csv")) 
```

# Cleaning

There are a few columns that we do not need or that contain NaN. So let's remove those.

#### Remove NAs

```{r}
dust<- dust %>%
  filter(!date %in% as.Date('2022-07-04'))
```

#### Select Dates for Daily Time series

```{r}
dust<-
  dust %>%
  filter(date %in% as.Date('2022-07-01') : as.Date('2022-07-31'))
```

### Plot T7sum over Daily Time series

<!--# Make this graph look nicer -->

```{r}
dust %>% ggplot()+ geom_line(
  aes(x = date,
      y = t7sum))+
  labs(title = "Dust Concentration over Daily Time Series",
       x = "Date",
       y = "Dust Concentration (AOD)")+
  theme_bw()

ggsave(here("results", "plots", "dust_july.png"))
```