### 2023-05-06 products/manuscript/supplement/Supplementary_Material.qmd

---
title: "Greenslit Project Supplementary Material"

subtitle: ""
author: Nathan Greenslit
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Additional Figures

## Dust for July

```{r}
knitr::include_graphics(here("results", "plots", "dust_july.png"))
```

**Dust AOD for month of July. Note slightly elevated dust levels prior to sampling period on the 5th. This may have impacted initial Vibrio levels on the 7th.**

## Distribution Tests (Q-Q Plots)

#### Copies per mL

![](images/paste-2F0DEEC8.png)

![](images/paste-0D0E739C.png)

![](images/paste-879F1B24.png)

#### Dust

![](images/paste-8CCC7BA6.png)

## Cross Correlation Analysis

![](images/paste-3B108E2C.png)

![![](images/paste-62178137.png)](images/paste-022F00A9.png)

**A negative Lag value indicates that x is a predictor of y, or in this case, dust is a predictor of copies_mL- so we will be looking at these negative values. A positive ACF indicates a positive relationship between the two variables, where a negative ACF indicates a negative relationship. According to the ccf plots, BO has a significant lag relationship a value of -1. In other words, dust occurs, and one day later, we see a response in copies_mL. C2 has a lag at -2, and RD at -1**